---
title: "BeSoccer Injury Web Scraping Process"
author: "Abdullah Zubair"
date: "2022-10-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We attempt at a web scraping of a page, BeSoccer to get the list of injuries of Gabriel Jesus. The challenge is to analyze the pattern of the URLs and later to turn the output into a tidy dataframe. We use the R packages listed below.

It is evidently clear that there may be a much quicker way to code this web scraping but I intended it to be flexible to any changes in Data such as other player's injury record.

## 1.0 Libraries

```{r message=FALSE,warning=FALSE}
library(rvest)
library(Rcrawler)
library(dplyr)
library(stringr)
library(lubridate)
library(tidyr)
```

## 2.0 Get Team Page

```{r}
leaguepath = "https://www.besoccer.com/competition/table/premier_league/2023"
teamlinks = LinkExtractor(leaguepath)[["InternalLinks"]]
teampages = grep("www.besoccer.com/team",teamlinks)
teampages = teamlinks[teampages]
# Remove the unwanted link
teampages = teampages[-1]
teampages
```

## 3.0 Generate Squad Page

-  Sample Page : https://www.besoccer.com/team/squad/afc-bournemouth
-  Just need to add /squad/ to the urls rather than do another round of LinkExtractor

```{r}
a = unlist(str_split(teampages,"https://www.besoccer.com/team"))
a = a[a != ""]
a = paste0("https://www.besoccer.com/team/squad",a)
head(a)
```

## 4.0 Get Player List

```{r}
players = LinkExtractor(a[1])[["InternalLinks"]]
playerspages = grep("www.besoccer.com/player",players)
playerspages = players[playerspages]
# Remove unwanted link
playerspages = playerspages[-1]
playerspages
```

## 5.0 Get the links to the page that has injury lists

-  Sample Page : https://www.besoccer.com/player/career-path/a-ramsdale-325210
-  Similarly like before, we just need to add the /career-path/ to the links

```{r}
a = unlist(str_split(playerspages,"https://www.besoccer.com/player/"))
a = a[a != ""]
player_cpath = paste0("https://www.besoccer.com/player/career-path/",a)
```

## 6.0 Extract Information

Now we access Gabriel Jesus' page at BeSoccer and start to extract information.

The URL is **`r player_cpath[21]`**

### 6.1 Create Club Table

On the webpage, we extracted his transfer history to specify which club he was injured for whilst playing.

```{r}
# Read URL
a1 = read_html(player_cpath[21]) 

# Get Club Data
cur_club = a1 %>% html_nodes("div#mod_detail_player_transfers.grid-item") %>% html_table() %>% as.data.frame()
colnames(cur_club) = c("date","from_club","to_club","fee")
cur_club = cur_club %>% mutate(date = str_squish(date),
                               from_club = str_squish(from_club),
                               to_club = str_squish(to_club),
                               fee = str_squish(fee)) %>% 
  separate(date, into = c("season","date"),sep = " ") %>% mutate(date = as.Date(date,format = "%d-%m-%Y")) 

# Rearranging the Data to have his club, starting and ending date at the club
start_club = vector()
end_club = vector()
startdate = vector()
enddate = vector()
for (i in 1:nrow(cur_club)) {
  # get club
  pat_loc_s = data.frame(str_locate_all(cur_club$from_club[i],"\\}"))
  pat_loc_e = data.frame(str_locate_all(cur_club$to_club[i],"\\}"))
  start_club[i] = str_sub(cur_club$from_club[i],start = pat_loc_s$end[1]+1)
  end_club[i] = str_sub(cur_club$to_club[i],start = pat_loc_e$end[1]+1)
  if(i == 1){
  startdate[i] = cur_club$date[i] 
  enddate[i] = Sys.Date() # just a date placeholder for current club's end date
}else if(i == nrow(cur_club)){
  startdate[i] = cur_club$date[i]-2 # just a placeholder for first club's start date
  enddate[i] = cur_club$date[i]
}else{
  startdate[i] = cur_club$date[i]
  enddate[i] = cur_club$date[i-1]-1
}
 }

cur_club = cur_club %>% mutate(from_club = str_squish(start_club),to_club = str_squish(end_club),
                                start_date = as.Date(startdate, origin = "1970-01-01"),
                                end_date = as.Date(enddate, origin = "1970-01-01")) %>% 
  select(club = to_club,start_date,end_date)

DT::datatable(cur_club)
```

### 6.2 Get Injury Data

Now that his employers are sorted, we move towards getting the actual injury data. This is how the Raw data looks like after removing whitespaces:

```{r}
# Get the injury data
pname = a1 %>% html_nodes("h2.title.ta-c") %>% html_text()

a2 = a1 %>% html_nodes("div#mod_detail_player_injuries.grid-item") %>% html_text()
# Remove Whitespace and special characters
a2 = str_squish(a2)

print(a2)
```

If we look closely, the Data is split by Year, similar how it is on the web page. We separate the data by Year so that it will be easier to tag the date of injury

```{r}
a3 = unlist(strsplit(a2,"Archive"))
a3 = str_squish(a3)
a3 = a3[a3 != ""]
head(a3)
```

## 7.0 Organizing into Data

There strategy is similar to previous method. Find a pattern and break it down into smaller pieces to fit into our Data frame.

```{r}
# Prepare pattern for Regex Later (To Extract the Injury Date)
month_name = month.name
month_name = paste(month_name,collapse = "|")

# Get the Year
inj_year = grep("[[:digit:]]{4}",a3[1],value = T)
inj_year = unlist(str_extract_all(a3,"(1|2)\\d{3}"))
```


```{r}
# Get the cause of injury
full_inj_df = data.frame()
for(j in 1:length(inj_year)){
  # Remove all spaces
  a4 = str_squish(a3[j])
  # Separate the cause and injury date
  cause_locator = data.frame(str_locate_all(a4,"Cause"))
  from_locator = data.frame(str_locate_all(a4,"From"))
  month_locator = data.frame(str_locate_all(a4,month_name))
  # Extract the Text
  inj_cause = vector()
  inj_date = vector()
  for (i in 1:nrow(cause_locator)) {
    inj_cause[i] = substr(a4,start = (cause_locator$end[i]+3),stop = (from_locator$start[i]-1)) 
    inj_date[i] = substr(a4,start = (from_locator$end[i]+2),stop = month_locator$end[(i*2)])
  }
  # Put in a Dataframe
  inj_df = data.frame("cause" = inj_cause,"date" = inj_date)
  inj_df = inj_df %>% mutate(year = inj_year[j])
  full_inj_df = bind_rows(full_inj_df,inj_df)
}
# Tidy the Data
full_inj_df = full_inj_df %>% separate(col = "date",into = c("date_start","date_end"),sep = "until") %>% 
  mutate(date_start = trimws(date_start),
         date_end = trimws(date_end))
# 

DT::datatable(full_inj_df)
```

If you monitor closely, there will be a duplicate Data whenever an absence starts from the current year and continues to the year after. For example, row 8 and 9.

We convert it into temporary data format and simply check whether the `month_end` is smaller than `month_start`. Afterwards, we converted the date into R's Date Class

```{r}
# Temporarily converting into date format for easier check
full_inj_df$date_start_temp = lubridate::month(as.Date(full_inj_df$date_start,format = "%d %B"))
full_inj_df$date_end_temp = lubridate::month(as.Date(full_inj_df$date_end,format = "%d %B"))
# Tag the duplicates
full_inj_df = full_inj_df %>% mutate(duplicate_marker = ifelse(date_end_temp < date_start_temp,"check","ok"))
# Removing duplicates 
t1 = full_inj_df %>% filter(duplicate_marker == "check")
toDelete <- seq(1, nrow(t1), 2)
t1 = t1[-toDelete,]
# binding it back together
full_inj_df = full_inj_df %>% filter(duplicate_marker != "check") %>% bind_rows(.,t1)
# Create a column stating year end
full_inj_df = full_inj_df %>% mutate(year_end = ifelse(duplicate_marker == "check",as.integer(year)+1,year),
                                       date_start = dmy(paste(date_start,year)),
                                       date_end = dmy(paste(date_end,year_end))
                                       ) %>% 
  arrange(desc(date_start))

# Keep necessary columns
full_inj_df = full_inj_df[,1:4]
DT::datatable(full_inj_df)
```

## 8.0 Adding Metadata

Last but not least, we added the club from the `club table` we retrieved previously. Then we add the Player's name into the table.

```{r}
# Matching the injury date with the club he was currently at
full_inj_df$club_at = NA
for (i in 1:nrow(cur_club)) {
  interval <- interval(cur_club$start_date[i], cur_club$end_date[i])
  index <-which(as.POSIXlt(full_inj_df$date_start) %within% interval)
  full_inj_df$club_at[index] <- cur_club$club[i]
}


# Complete the table with the player name
full_inj_df = full_inj_df %>% mutate(player_name = pname) %>% relocate(c(player_name,club_at),.before = cause)
DT::datatable(full_inj_df)
```



